Visual Diagnostics for Constrained Optimisation with Application to Guided Tours

### 2021-03-18 paper/zhang-cook-laa-langrene-menendez.Rmd
interesting index polished

The focus of this paper is on the optimisation problem arising in the projection pursuit guided tour [@buja2005computational], an exploratory data analysis tool used for detecting interesting structures in high-dimensional data through a set of lower-dimensional projections [@cook2008grand]. The goal of the optimisation is to identify the projection, characterised by the projection basis, that gives the most interesting low-dimensional view. A view is interesting if it can show some unexpected structure of the data, e.g., bi-modality, clustering and outliers.

@@

Let $\mathbf{X}_{n \times p}$ be the data matrix with $n$ observations in $p$ dimensions. A $d$-dimensional projection is a linear transformation from $\mathbb{R}^p$ into $\mathbb{R}^d$ defined as $\mathbf{Y} = \mathbf{X} \cdot \mathbf{A}$, where $\mathbf{Y}_{n \times d}$ is the projected data and $\mathbf{A}_{p\times d}$ is the projection basis matrix. We define $f: \mathbb{R}^{n \times d} \mapsto \mathbb{R}$ to be an index function that maps the projected data $\mathbf{Y}$ onto a scalar value. This is commonly known as the projection pursuit index function, or just index function, and is used to measure the "interestingness" of a given projection. An interesting projection shows structures that are non-normal and this has been argued for two reasons: Theoretical proofs from @diaconis1984asymptotics show that projections tend to be normal as $n$ and $p$ approaching infinity under certain conditions. @huber1985projection illustrates that, for indices based on entropy measures, [...]. Index functions  proposed in the litearture include early indices that can be categorised by measuring the $L^2$ distance between the projection and a normal distribution.(Legendre index [@friedman1974projection], Hermite index [@hall1989polynomial], and natural Hermite index [@cook1993projection]); chi-square index [@posse1995projection] for detecting spiral structure; kurtosis [@Loperfido2020] and skewness [@Loperfido2018] indices for detecting outliers in financial time series; and most recently, scagnostic indices [@laa2020using] for summarising structures in scatterplot matrices based on eight scagnostic measures.

>>>>

The focus of this paper is on the optimisation problem arising in the projection pursuit guided tour [@buja2005computational], an exploratory data analysis tool used for detecting interesting structures in high-dimensional data through a set of lower-dimensional projections [@cook2008grand]. The goal of the optimisation is to identify the projection, characterised by the projection basis, that gives the most interesting low-dimensional view. A view is interesting if it can show some structures of the data, e.g., bi-modality, clustering and outliers, other than normal distribution.

@@

Let $\mathbf{X}_{n \times p}$ be the data matrix with $n$ observations in $p$ dimensions. A $d$-dimensional projection is a linear transformation from $\mathbb{R}^p$ into $\mathbb{R}^d$ defined as $\mathbf{Y} = \mathbf{X} \cdot \mathbf{A}$, where $\mathbf{Y}_{n \times d}$ is the projected data and $\mathbf{A}_{p\times d}$ is the projection basis matrix. We define $f: \mathbb{R}^{n \times d} \mapsto \mathbb{R}$ to be an index function that maps the projected data $\mathbf{Y}$ onto a scalar value. This is commonly known as the projection pursuit index function, or just index function, and is used to measure the "interestingness" of a given projection. An interesting projection shows structures that are non-normal and this has been argued for two reasons: Theoretical proofs from @diaconis1984asymptotics show that projections tend to be normal, as $n$ and $p$ approaching infinity under certain conditions. Index functions proposed in the litearture include early indices that can be categorised by measuring the $L^2$ distance between the projection and a normal distribution: Legendre index [@friedman1974projection], Hermite index [@hall1989polynomial], and natural Hermite index [@cook1993projection]; chi-square index [@posse1995projection] for detecting spiral structure; kurtosis index [@Loperfido2020] and skewness index [@Loperfido2018] for detecting outliers in financial time series; and most recently, scagnostic indices [@laa2020using] for summarising structures in scatterplot matrices based on eight scagnostic measures.

### 2021-03-18 paper/zhang-cook-laa-langrene-menendez.tex
interesting index polished

The focus of this paper is on the optimisation problem arising in the
projection pursuit guided tour \citep{buja2005computational}, an
exploratory data analysis tool used for detecting interesting structures
in high-dimensional data through a set of lower-dimensional projections
\citep{cook2008grand}. The goal of the optimisation is to identify the
projection, characterised by the projection basis, that gives the most
interesting low-dimensional view. A view is interesting if it can show
some unexpected structure of the data, e.g., bi-modality, clustering and
outliers.

@@

Let \(\mathbf{X}_{n \times p}\) be the data matrix with \(n\)
observations in \(p\) dimensions. A \(d\)-dimensional projection is a
linear transformation from \(\mathbb{R}^p\) into \(\mathbb{R}^d\)
defined as \(\mathbf{Y} = \mathbf{X} \cdot \mathbf{A}\), where
\(\mathbf{Y}_{n \times d}\) is the projected data and
\(\mathbf{A}_{p\times d}\) is the projection basis matrix. We define
\(f: \mathbb{R}^{n \times d} \mapsto \mathbb{R}\) to be an index
function that maps the projected data \(\mathbf{Y}\) onto a scalar
value. This is commonly known as the projection pursuit index function,
or just index function, and is used to measure the ``interestingness''
of a given projection. An interesting projection shows structures that
are non-normal and this has been argued for two reasons: Theoretical
proofs from \citet{diaconis1984asymptotics} show that projections tend
to be normal as \(n\) and \(p\) approaching infinity under certain
conditions. \citet{huber1985projection} illustrates that, for indices
based on entropy measures, {[}\ldots{}{]}. Index functions proposed in
the litearture include early indices that can be categorised by
measuring the \(L^2\) distance between the projection and a normal
distribution.(Legendre index \citep{friedman1974projection}, Hermite
index \citep{hall1989polynomial}, and natural Hermite index
\citep{cook1993projection}); chi-square index
\citep{posse1995projection} for detecting spiral structure; kurtosis
\citep{Loperfido2020} and skewness \citep{Loperfido2018} indices for
detecting outliers in financial time series; and most recently,
scagnostic indices \citep{laa2020using} for summarising structures in
scatterplot matrices based on eight scagnostic measures.

>>>>

The focus of this paper is on the optimisation problem arising in the
projection pursuit guided tour \citep{buja2005computational}, an
exploratory data analysis tool used for detecting interesting structures
in high-dimensional data through a set of lower-dimensional projections
\citep{cook2008grand}. The goal of the optimisation is to identify the
projection, characterised by the projection basis, that gives the most
interesting low-dimensional view. A view is interesting if it can show
some structures of the data, e.g., bi-modality, clustering and outliers,
other than normal distribution.

@@

Let \(\mathbf{X}_{n \times p}\) be the data matrix with \(n\)
observations in \(p\) dimensions. A \(d\)-dimensional projection is a
linear transformation from \(\mathbb{R}^p\) into \(\mathbb{R}^d\)
defined as \(\mathbf{Y} = \mathbf{X} \cdot \mathbf{A}\), where
\(\mathbf{Y}_{n \times d}\) is the projected data and
\(\mathbf{A}_{p\times d}\) is the projection basis matrix. We define
\(f: \mathbb{R}^{n \times d} \mapsto \mathbb{R}\) to be an index
function that maps the projected data \(\mathbf{Y}\) onto a scalar
value. This is commonly known as the projection pursuit index function,
or just index function, and is used to measure the ``interestingness''
of a given projection. An interesting projection shows structures that
are non-normal and this has been argued for two reasons: Theoretical
proofs from \citet{diaconis1984asymptotics} show that projections tend
to be normal, as \(n\) and \(p\) approaching infinity under certain
conditions. Index functions proposed in the litearture include early
indices that can be categorised by measuring the \(L^2\) distance
between the projection and a normal distribution: Legendre index
\citep{friedman1974projection}, Hermite index
\citep{hall1989polynomial}, and natural Hermite index
\citep{cook1993projection}; chi-square index \citep{posse1995projection}
for detecting spiral structure; kurtosis index \citep{Loperfido2020} and
skewness index \citep{Loperfido2018} for detecting outliers in financial
time series; and most recently, scagnostic indices \citep{laa2020using}
for summarising structures in scatterplot matrices based on eight
scagnostic measures.
